---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

# faviconPlease

[![R-CMD-check](https://github.com/jdblischak/faviconPlease/workflows/R-CMD-check/badge.svg)](https://github.com/jdblischak/faviconPlease/actions)

```{r description, results='asis', echo=FALSE}
cat(read.dcf("DESCRIPTION", fields = "Description"))
```

```{r example}
library(faviconPlease)
faviconPlease("https://github.com/")
```

## Installation

```{r installation, eval=FALSE}
install.packages("remotes")
remotes::install_github("jdblischak/faviconPlease")
```

## Code of Conduct
  
Please note that the faviconPlease project is released with a [Contributor Code
of Conduct](https://contributor-covenant.org/version/2/0/CODE_OF_CONDUCT.html).
By contributing to this project, you agree to abide by its terms.

## Default strategy

By default, `faviconPlease()` uses the following strategy to find the URL to the
favicon for a given website. It stops once it finds a URL and returns it.

1. Download the HTML file and search its `<head>` for any `<link>` elements with
`rel="icon"` or `rel="shortcut icon"`.

1. Download the HTML file at the root of the server (i.e. discard the path) and
search its `<head>` for any `<link>` elements with `rel="icon"` or
`rel="shortcut icon"`.

1. Attempt to download a file called `favicon.ico` at the root of the server.
This is the default location that a browser looks if the HTML file does not
specify an alternative location in a `<link>` element. If the file `favicon.ico`
is successfully downloaded, then this URL is returned.

1. If the above steps fail, as a fallback, use the [favicon service][ddg-icon]
provided by the search engine [DuckDuckGo][ddg], e.g.
https://icons.duckduckgo.com/ip3/github.com.ico. This provides a nice default
for websites that don't have a favicon (or can't be easily found), e.g.
https://icons.duckduckgo.com/ip3/abc.ico

[ddg]: https://duckduckgo.com/
[ddg-icon]: https://help.duckduckgo.com/duckduckgo-help-pages/features/favicons/

## Extending faviconPlease


## Troubleshooting

